### 2020-05-06 Case Studies/03_spam_identification.rmd
spam case study clean-up
spam case study clean-up
spam identification, recursive partitioning approach
spam identification, recursive partitioning approach

@@

CTloc <- sapply(headerList, grep, pattern = "Content-Type")
CTloc

@@

   invisible(msgWordsList)
}
```

@@

isSpam <- rep(c(FALSE, FALSE, FALSE, TRUE, TRUE), numMsgs)

@@

```{r}
numEmail <- length(isSpam)

@@

log(spamWordsProbs) - log(hamWordsProbs)

@@

```{r}
sampleStruct <- emailStruct[ 1 ]
```

@@

@@

@@

>>>>

library(rpart, quietly = TRUE, warn.conflicts = FALSE)
library(rpart.plot, quietly = TRUE, warn.conflicts = FALSE)

@@

CTloc <- sapply(headerList, grep, pattern = "Content-Type")

@@

   invisible(msgWordsList)
}

@@

isSpam <- c(rep(FALSE, numMsgs[1]), 
            rep(FALSE, numMsgs[2]), 
            rep(FALSE, numMsgs[3]), 
            rep(TRUE, numMsgs[4]),
            rep(TRUE, numMsgs[5]))

@@

numEmail <- length(isSpam)

@@

probs <- log(spamWordsProbs) - log(hamWordsProbs)

head(probs)
```

@@

```{r}
sampleStruct <- emailStruct[ 1:15 ]
```

@@

```{r}
header <- sampleStruct[[1]]$header
subject <- header["Subject"]
els <- strsplit(subject, "")
all(els %in% LETTERS)

testSubject <- c("DEAR MADAM", "WINNER!", "")

els <- strsplit(testSubject, "")
sapply(els, function(subject) all(subject %in% LETTERS))

gsub("[[:punct:] ]", "", testSubject)
gsub("[^[:alpha:]]", "", testSubject)

```{r}
isYelling <- function(msg) {
   if( "Subject" %in% names(msg$header) ) {
      el <- gsub("[^[:alpha:]]", "", msg$header["Subject"])

      if ( nchar(el) > 0 )
         nchar(gsub("[A-Z]", "", el) < 1 )
      else
         FALSE
   } else {
      NA
   }
}

perCaps <- function(msg) {

   body <- paste(msg$body, collapse = "")

   # Return NA if the body of the message is "empty"
   if(length(body) == 0 || nchar(body) == 0) return (NA)

   # Eliminate non-alpha characters
   body <- gsub("[^[:alpha:]]", "", body)
   capText <- gsub("[^A-Z]", "", body)
   100 * nchar(capText)/nchar(body)
}

```{r}
sapply(sampleStruct, perCaps)
```

@@

funcList <- list(

   isRe = function(msg) {
      "Subject" %in% names(msg$header) &&
         length(grep("^[ ]*Re:", msg$header[["Subject"]])) > 0
   },
   numLines = function(msg) {
      length(msg$body)
   },
   isYelling = function(msg) {
      if( "Subject" %in% names(msg$header) ) {
         el <- gsub("[^[:alpha:]]", "", msg$header["Subject"])

         if ( nchar(el) > 0 )
            nchar(gsub("[A-Z]", "", el) < 1 )
         else
            FALSE
      } else {
         NA
      }
   },
   perCaps = function(msg) {

      body <- paste(msg$body, collapse = "")

      # Return NA if the body of the message is "empty"
      if(length(body) == 0 || nchar(body) == 0) return (NA)

      # Eliminate non-alpha characters
      body <- gsub("[^[:alpha:]]", "", body)
      capText <- gsub("[^A-Z]", "", body)
      100 * nchar(capText)/nchar(body)
   }
)

@@

```{r}
lapply(funcList, function(func)
   sapply(sampleStruct, function(msg) func(msg)))
```

```{r}
createDerivedF <- function(email = emailStruct, operations = funcList,
                           verbose = F)
{
   els <- lapply(names(operations),
                 function(id) {
                     if(verbose) print(id)
                     e <- operations[[id]]
                     v <- if(is.function(e))
                           sapply(email, e)
                        else
                           sapply(email, function(msg) eval(e))
                     v
                  })

   df <- as.data.frame(els)
   names(df) <- names(operations)

   invisible(df)  
}
```

```{r}
sampleDF <- createDerivedF(sampleStruct)

```{r}
spam_data <- file.path(data.dir, "spamAssassinDerivedDF.rda")

load(spam_data)
```

```{r}
perCaps2 <- function(msg) {

   body <- paste(msg$body, collapse = "")

   # return NA if the body of the message is "empty"
   if(length(body) == 0 || nchar(body) == 0) return(NA)

   # eliminate non-alpha characters and empty lines
   body <- gsub("[^[:alpha:]]", "", body)
   els <- unlist(strsplit(body, ""))
   ctCap <- sum(els %in% LETTERS)
   100 * ctCap / length(els)
}

```{r}
pC <- sapply(emailStruct, perCaps)
pC2 <- sapply(emailStruct, perCaps2)

identical(pC, pC2)
```

```{r}
indNA <- which(is.na(emailDF$subExcCt))

indNoSubject <- which(sapply(emailStruct,
                             function(msg)
                                !("Subject" %in% names(msg$header))))

all(indNA == indNoSubject)
```

```{r}
all(emailDF$bodyCharCt > emailDF$numLines)
```

long_lines <- head(sort(emailDF$numLines, decreasing = T), 10)

rem <- which(emailDF$numLines %in% long_lines)

ggplot(emailDF[-rem, ], aes(bodyCharCt, numLines)) +
   geom_point(aes(col = bodyCharCt)) +
   geom_smooth(method = "lm") +
   scale_x_continuous(lim = c(0, 35000))

#### Exploring the email Feature Set

```{r}
percent <- emailDF$perCaps
isSpamLabs <- factor(emailDF$isSpam, labels = c("ham", "spam"))
boxplot(log(1 + percent) ~ isSpamLabs,
        ylab = "Percent Capitals (log)")

ggplot(emailDF, aes(perCaps, bodyCharCt, col = isSpam)) +
   geom_point() +
   scale_y_log10() +
   scale_x_log10()

colI <- c("#4DAF4A80", "#984EA380")
logBodyCharCt <- log(1 + emailDF$bodyCharCt)
logPerCaps <- log(1 + emailDF$perCaps)
plot(logPerCaps ~ logBodyCharCt, xlab = "Total Characters (log)",
     ylab = "Percent Capitals (log)",
     col = colI[1 + emailDF$isSpam],
     xlim = c(2, 12), pch = 19, cex = 0.5)
```

```{r}
table(emailDF$numAtt, isSpamLabs)
```

```{r}
colM <- c("#E41A1C80", "#377EB880")
isRe <- factor(emailDF$isRe, labels = c("no Re:", "Re:"))
mosaicplot(table(isSpamLabs, isRe), main = "",
           xlab = "", ylab = "", color = colM)

fromNE <- factor(emailDF$numEnd, labels = c("No #", "#"))
mosaicplot(table(isSpamLabs, fromNE), color = colM,
           main = "", xlab = "", ylab = "")
```

### Fitting Recursive Partition

```{r}
setupRpart <- function(data) {
   logicalVars <- which(sapply(data, is.logical))
   facVars <- lapply(data[, logicalVars],
                     function(x) {
                        x = as.factor(x)
                        levels(x) = c("F", "T")
                        x
                     })
   cbind(facVars, data[, - logicalVars])
}

emailDFrp <- setupRpart(emailDF)
```

```{r}
set.seed(418910)

testDF <- rbind( emailDFrp[ emailDFrp$isSpam == "T", ][testSpamIdx, ],
                 emailDFrp[ emailDFrp$isSpam == "F", ][testHamIdx, ])

trainDF <- rbind( emailDFrp[ emailDFrp$isSpam == "T", ][-testSpamIdx, ],
                 emailDFrp[ emailDFrp$isSpam == "F", ][-testHamIdx, ])

```{r}
rpartFit <- rpart(isSpam ~ ., data = trainDF, method = "class")

prp(rpartFit, extra = 1)
```

```{r}
predictions <- predict(rpartFit,
                       newdata = testDF[, names(testDF) != "isSpam"],
                       type = "class")

predsForHam <- predictions[ testDF$isSpam == "F" ]
summary(predsForHam)

sum(predsForHam == "T", na.rm = T) / length(predsForHam)

predsForSpam <- predictions[ testDF$isSpam == "T" ]
sum(predsForSpam == "F", na.rm = T) / length(predsForSpam)
```

```{r}
args(rpart.control)
```

```{r}
complexityVals <- c(seq(0.00001, 0.0001, length = 19),
                    seq(0.0001, 0.001, length = 19),
                    seq(0.001, 0.005, length = 9),
                    seq(0.005, 0.01, length = 9))

fits <- lapply(complexityVals, function(x) {
   rpartObj <- rpart(isSpam ~ ., data = trainDF,
                     method = "class",
                     control = rpart.control(cp=x))

   predict(rpartObj,
           newdata = testDF[, names(testDF) != "isSpam"],
           type = "class")
})

```{r}
spam <- testDF$isSpam == "T"
numSpam <- sum(spam, na.rm = T)
numHam <- sum(!spam, na.rm = T)

errs <- sapply(fits, function(preds) {
   typeI = sum( preds[ !spam ] == "T", na.rm = T) / numHam
   typeII = sum( preds[ spam ] == "F", na.rm = T) / numSpam
   c(typeI = typeI, typeII = typeII )
})

errs

err_df <- data.table(t(errs))
